---
title: "Motor Trend Project - Regression Coursera - JHU Course"
author: "Gito"
date: "2024-11-02"
output: 
  word_document: default
---


# EXECUTIVE SUMMARIES

This study is part of Coursera Course on Regression Model.
We will use R package mtcars
You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

“Is an automatic or manual transmission better for MPG”

"Quantify the MPG difference between automatic and manual transmissions"  

Analytical Process :

1. Data Processing & Variable Explanation
2. Confidence level test
3. Model Selection
4. Summary Conclusion
5. Appendixes (Pairs plot and Residuals Plot)


# Data Processing and Variable Explanation
### Data Processing
We load mtcars and do some data processing and explanation. 'mtcars' data is part of datasets library and we ll load necessary library like, ggplot2, tidyverse,dplyr, GGally.
Below variables are taken from the r studio help, ?mtcars
We ll change variable class for listed below to factor

1. mpg	Miles/(US) gallon --> outcome
2. cyl	Number of cylinders --> need to be factor variable
3. disp	Displacement (cu.in.) 
4. hp	Gross horsepower
5. drat	Rear axle ratio
6. wt	Weight (1000 lbs)
7. qsec	1/4 mile time 
8. vs	Engine (0 = V-shaped, 1 = straight) --> need to be factor variable
9. am	Transmission (0 = automatic, 1 = manual) --> need to be factor variable
10.	gear	Number of forward gears --> need to be factor variable
11.	carb	Number of carburetors --> need to be factor variable


```{r }
library(datasets)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(GGally)
# head(mtcars)
raw_mtcars <- mtcars
raw_mtcars[,c(2,8,9,10,11)] <- lapply(raw_mtcars[,c(2,8,9,10,11)],as.factor)
levels(raw_mtcars$am) <- c("AT","MN")
```

### Variable Explanation and Corelation 

GGPairs plot showing corelation plot for each variable to understand which variable are corelated so we can remove it and we will explain it on lm model

1. Variable am and vs seems correlated, similar impact to mpg --> we will use am

2. Variable am and cyc seems uncorelated, it has negative impact to mpg --> we will use cyc variable as confound variable

3. Variable am and disp seems uncorelated, it has different in distribution

4. Variable am and drat seems uncorelated, it has different in distribution

5. Variable am and wt seems uncorelated, it has different in distribution

6. Variable am and qseq seems corelated, it has similar distributions

7. Variable am and hp seems uncorelated

We will see this distribution to get better understanding

```{r }
head(raw_mtcars)
glimpse(raw_mtcars)
ggplot(raw_mtcars,aes(x=am,y=mpg))+geom_boxplot(aes(fill = am))+theme_bw()+labs(x="Auto or Manual",y="Miles per gallon",title = "Miles per Gallons vs Auto/Manual")
```

Below are Correlation Density for above variable
We will see mpg distribution per transmissions to get idea how data distributed.
We can see mpg vs AM are not perfectly separated, there will be some unknown factor to get perfect modelling
```{r }
ggplot(raw_mtcars,aes(mpg))+geom_density(aes(fill = am),alpha = 0.3, width = 0.3)+labs(x="mpg",y="Density",title = "MPG vs AM Density")
```
Below are variable corelation density
We see here that 
1. DISP and AM are unclearly separated, this suggest that DISP is abit corelated with AM, this suggest not to use DISP as regressor

2. DRAT and AM are separated, this suggest that DRAT is uncorelated with AM, this suggest to use DRAT as regressor

3. WT and AM are separated, this suggest that WT is uncorelated with AM, this suggest to use WT as regressor

4. QSEC and AM are not separated, this suggest that QSEC is corelated with AM, this suggest not to use DISP as regressor

5. HP and AM are separated, this suggest that HP is uncorelated with AM, this suggest to use HP as regressor

6. We can see cyl is also uncorelated based on pairs on appendix

```{r figures-side, out.width = '50%'}
ggplot(raw_mtcars,aes(disp))+geom_density(aes(fill = am),alpha = 0.3)+labs(x="disp",y="Density",title = "DISP vs AM Density")
ggplot(raw_mtcars,aes(drat))+geom_density(aes(fill = am),alpha = 0.3)+labs(x="drat",y="Density",title = "DRAT vs AM Density")
ggplot(raw_mtcars,aes(wt))+geom_density(aes(fill = am),alpha = 0.3)+labs(x="wt",y="Density",title = "WT vs AM Density")
ggplot(raw_mtcars,aes(qsec))+geom_density(aes(fill = am),alpha = 0.3)+labs(x="qsec",y="Density",title = "QSEC vs AM Density")
ggplot(raw_mtcars,aes(hp))+geom_density(aes(fill = am),alpha = 0.3)+labs(x="hp",y="Density",title = "HP vs AM Density")
```


# Confidence Level
This is to calculate the statistical inference about the MPG difference for Automatic and Manual Transmission. From previous box plot, we can see clear MPG difference  for Automatic and Manual Transmission.

```{r }
t.test(filter(raw_mtcars,am == "AT")$mpg, filter(raw_mtcars,am == "MN")$mpg)
```
We can see the t value = -3.77 and confidence interval is negative value with no 0 crossing, it suggests that MPG value for Automatic and Manual Transmission is different with Manual Transmission has higher Miles per Gallon.
P-Value of 0.1% is sufficient to reject Null Hipothesis and suggested Alternative Hipothesis

# Modelling
We will use multiple modelling to see which model have better understanding 
### 1. AM only as regressor
```{r }
model1 <- lm(mpg~am,raw_mtcars)
summary(model1)
```
### 2. All Variable as regressor
```{r }
model2 <- lm(mpg~cyl+disp+hp+drat+wt+qsec+vs+am+gear+carb,raw_mtcars)
summary(model2)
```
### 3. Only uncorelated variable as regressor (DRAT,HP,WT,CYL and AM)
Below Linear Model have better R-Squared error compare to previous 2 models. This model suggest below points :
1. Each unit change on Manual Transmission Type has ~1.7mpg difference
2. Each unit change on Weight has -2.4mpg difference
3. Each unit change on Rear Axle Ratio has ~0.27mpg difference
4. Each unit change on Horse Power has ~-0.03mpg difference
5. Each unit change on Cylinder has ~-2.9 and -1.9 mpg difference for Cyl 6&8 respectively

```{r }
model3 <- lm(mpg~cyl+hp+drat+wt+am,raw_mtcars)
summary(model3)
```

# Summary and Conclusion
As per Linear Regression Modelling we can conclude below points :

1. Transmission Type suggest to have significant impact on Miles per Gallons based on T - test statistic suggest that it has significant difference

2. Here are factors impacting MPG based on linear model, Transmission Type, Weight, Rear Axle, Horse Power, Cylinder

3. Model suggest that Manual Transmission has better MPG

4. Some unexplainable factors due to Automatic and Manual Transmission MPG Density data are not perfectly segregated 



# APENDIXES
Correlation Pairs Diagram

```{r out.width = '50%'}
ggpairs(raw_mtcars,mapping = aes(col = am)) + theme_bw()
```

Residual diagram of model 1
```{r out.width = '50%'}
plot(model1)
```
Residual plot of model 2
```{r out.width = '50%' }
plot(model2)
```

residual plot of model 3 (best model)
```{r out.width = '50%'}
plot(model3)
```